\documentclass[11pt]{article}

\usepackage{setspace}
\usepackage[letterpaper,margin=1in]{geometry}
\usepackage[parfill]{parskip}
\usepackage{graphicx}
\graphicspath{ {images/} }

\title{Specifications for RSA Efficiency Research}
\author{George Wood}
\date{Last Modified: 3/7/2018}

\begin{document}
\maketitle

\thispagestyle{empty}

\section{Misc. Hardware Info}

\begin{enumerate}
\item{Raspberry Pi
	\begin{itemize}
	\item{Use Raspbian OS}
	\item{There may be implementation issues with Raspbian.}
	\end{itemize}
	}
\item{Arduino
	\begin{itemize}
	\item{Will probably melt.}
	\end{itemize}
	}
\item{Build hardware list, priority on RAM, processor, OS, and possibly GPU.
	}
\end{enumerate}


\section{RSA Specs}

\begin{enumerate}
	\item{Don't worry about randomness - data will be pre-generated and consistent across trials.}
	\item{Don't worry about padding for same reason}
	\item{Use data of sizes: 64B, 1kB, 32kB, 64kB, 128kB, 1MB}
	\item{Use keys of size: 512b, 1024b, 2048b, 4096b}
\end{enumerate}

\section{Measurements}

\begin{enumerate}
\item{Power consumption
	\begin{itemize}
	\item{Kill-A-Watt for physical measurements}
	\item{Software such as MS's Joulemeter}
	\end{itemize}
	}
\item{Cache/RAM IO Speed (Transfers/second)
	\begin{itemize}
	\item{Determined largely by RAM, info released by manufacturers}
	\end{itemize}
	}
\item{Keep track of RAM usage and CPU usage}
	
\end{enumerate}



\section{To-Dos/Deadlines}

\begin{enumerate}
\item{Deadlines
	\begin{itemize}
	\item{Thursday, 3/8/2018: Kill-A-Watt arrives, configure and test.}
	\item{Sunday, 3/11/2018: Complete UI for performing tests/collecting data.}
	\item{Monday, 3/12/2018 - Tuesday, 3/13/2018: Perform tests and collect data on all available machines.}
	\item{Friday, 3/16/2018: Complete preliminary organization of all data collected up to this point.}
	\item{Friday, 3/30/2018 - Saturday, 3/31/2018: Perform tests and collect data on all remaining machines.}
	\item{Tuesday, 4/3/2018: Complete remaining organization of all data. Perform and record statistical analysis of data.}
	\item{Thursday, 4/5/2018: Complete first draft of paper. }
	\end{itemize}
	}
\end{enumerate}

\section{Etc}
\begin{enumerate}
\item{Other Papers
	\begin{itemize}
	\item{Wyatt Yost}
	\item{Thomas Graves}
	\end{itemize}
	}
\end{enumerate}



\end{document}